---
title: "Wine Qaulity Analysis"
author: "Narendra Bandi"
date: "08/04/2021"
output: pdf_document
fontsize: 12pt 
---
<style>
table {
  margin: auto;
  border-top: 1px solid #666;
  border-bottom: 1px solid #666;

}
table thead th { border-bottom: 1px solid #ddd; }
th, td { padding: 5px; }
thead, tfoot, tr:nth-child(even) { background: #eee; }
</style>


Wine quality prediction using 
Multinomial LOgistic regression, Decision Tree and Random Forest methods.
Quality is assigned discrete values from 1 to 10.


```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries, echo=T, include=FALSE, message=F}
library(readr)
library(caret)
library(dplyr)
library(corrplot)
library(MASS)
library(e1071)
```

##### Read data from file, the delimiter is ";"

```{r Load data, echo=T, message= F}
red_data <- read_delim("winequality-red.csv",delim  = ";")
dim(red_data)
```
```{r echo=TRUE, message=F}
knitr::kable(head(red_data, 10), align = "c")
```
##### Distribution of quality:
```{r echo= T, message= F}
quality.tbl <- table(red_data$quality)
knitr::kable(quality.tbl, align = "c")
barplot(quality.tbl,horiz = F,col = quality.tbl,  xlab = "Quality", main= "Quality distribution")

```

##### Linear correlation :
```{r echo=T, message=F}
corr_mtrx <- cor(red_data, method = "pearson")

corrplot(corr_mtrx, type = "upper",order = "AOE", 
         tl.col = "black", tl.srt = 90)
knitr::kable(round(corr_mtrx,2), align = "c")
```
```{r echo= T, message=F}
red_data$quality <-  factor(red_data$quality, levels=c(1:10), ordered=TRUE)
```

Split data into Train and Testing using stratification sampling

```{r echo= T, message=F}
row_idx <- createDataPartition(red_data$quality,p = 0.25,list = F)
train.df <- red_data[-row_idx[,1],]
validation.df <- red_data[row_idx[,1],]

```
##### Ordered  Multinomial Logistc Regression approach (quality is ordered)
```{r echo=T, message= F}
model_ord_log <- polr(quality ~ ., data = train.df, Hess=TRUE)
summary(model_ord_log)
```
###### predict the quality for test data.
```{r echo=T, message= F}
predicted_quality <- predict(model_ord_log, newdata = validation.df)
```

```{r echo= T , message = F}
confusion_mtrx <- confusionMatrix(predicted_quality, validation.df$quality)
confusion_mtrx
```

#### Decision Tree approach

```{r echo=T, message=F}
model.dt <- rpart::rpart(quality ~., data= train.df)
predicted.dt <- predict(model.dt, validation.df, type = "class")
confusion.dt <- confusionMatrix(predicted.dt, validation.df$quality)
confusion.dt
```
##### Random Forest approach


```{r echo=T}
train.col.df <- train.df
valid.col.df <- validation.df
col_names <- colnames(train.col.df)
colnames(train.col.df) <-   gsub(" ", "_", col_names)
colnames(valid.col.df) <- gsub(" " , "_", col_names)
 
```
```{r echo= T, message= FALSE}
library(randomForest)
train.col.df$quality <- droplevels(train.col.df$quality)
model.rf <- randomForest(quality ~., data= train.col.df, ntree=300, mtry=4, importance=TRUE)
```
```{r echo= T, message= F}

predicted.rf <- predict(model.rf, valid.col.df, type = "class")
predicted.prob.rf <- predict(model.rf, valid.col.df,type="prob")
confusion.rf <- confusionMatrix(predicted.rf, valid.col.df$quality)
confusion.rf
```



